Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity

### 2016-07-09 acknowledge.tex
implemented final comments from Naeem

The authors wish to express their gratitude to Art Frankel at the U.S.~Geological Survey (USGS) for providing access to the Smoothed Probabilistic Seismic Hazard Analysis software and to Morgan Moschetti, also at the USGS, for providing advise on the use of the software. The authors also thank Andrzej Kijko at the University of Pretoria for providing access to the software for computing the seismicity parameters; and Mehdi Zare at the International Institute of Earthquake Engineering and Seismology (IIEES) in Tehran for kindly providing the original catalog data and for his assistance in interpreting previous catalog information. Access to data provided by IIEES is also greatly appreciated. This research was possible thanks to support by the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunsent Laws (SB 1510 and HB 1608, 2015--2016).

>>>>

The authors wish to express their gratitude to Art Fran\-kel at the U.S.~Geological Survey (USGS) for providing access to the Smoothed Probabilistic Seismic Hazard Analysis software and to Morgan Moschetti, also at the USGS, for providing advise on the use of the software. The authors also thank Andrzej Kijko at the University of Pretoria for providing access to the software for computing the seismicity parameters; and Mehdi Zare at the International Institute of Earthquake Engineering and Seismology (IIEES) in Tehran for kindly providing the original catalog data and for his assistance in interpreting previous catalog information. Access to data provided by IIEES is also greatly appreciated. This research was possible thanks to support by the Center for Earthquake Research and Information (CERI) at the University of Memphis. CERI is designated as a Center of Excellence by the Tennessee Board of Regents and is funded in part by the State of Tennessee under State Sunsent Laws (SB 1510 and HB 1608, 2015--2016).

### 2016-07-09 completeness.tex
finalizing results section
implemented final comments from Naeem

The start-year of completeness is usually identified by inspection. A simple approach used by \citet{Frankel1995} and others is to select the year at which a plot of the cumulative number of events with respect to time becomes (or starts to resemble) a straight line. This is done for different magnitude ranges. Following this approach, we determined the year of completeness for magnitude ranges 3--4, 4--5, 5--6 and 6--7. Our choices are illustrated in Fig.~\ref{fig:completeness}, where we show the normalized cumulative number of events with time for each of the seismic zones and magnitude ranges, and indicate our selection of the year of completeness in each case. The values shown in this figure and those corresponding to the uniform model of the entire region of interest are summarized in Table \ref{tab:completeness}. For events $M_w>7$, in which case there are no sufficient data to determine a proper threshold, we assumed the catalog to be complete from the earliest time at which there is record of large magnitude earthquakes.

@@

We defined these year intervals based on the data magnitude-time distribution (see Fig.~\ref{fig:scatter}) and the idea of distinguishing between historical, early-instrumental, and modern-instrumental catalogs. Having augmented the original catalog with $M>3$ earthquakes beginning in 2000 also influenced our choice of the intervals because at this point the uniformity of the catalog is altered. No value of $M_c$ was computed for pre-historical (before 1800) data because of scarcity of data. Last, we also computed $M_c$ values for the complete regional catalogs, but in this case, in order to maintain uniformity throughout time, we only considered $M>4$ events. Fig.~\ref{fig:mc} illustrates the selection of $M_c$ for the case of the main three northern Iran seismic regions, and Table \ref{tab:completeness} presents the final selection of $M_c$ values for these and the other regions, including the uniform northern Iran model for the complete region of interest.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-06.pdf} 
    \caption{Selection of completeness magnitude ($M_c$) values for the three main seismic regions in northern Iran. The blue squares indicate the computed goodness-of-fit test values and the green circles indicate the cumulative number of events. Both are shown as functions of the earthquake magnitude. The former is plotted in linear scale (left axes) and the latter is plotted in logarithmic scale (right axes). The horizontal dashed lines indicate desired thresholds for the goodness-of-fit values at 80 and 90 percent. The vertical red lines show indicate the chosen $M_c$ values for each zone and catalog year interval.}
    \label{fig:mc}
\end{figure*}

>>>>

The start-year of completeness is commonly identified by inspection. A simple approach used by \citet{Frankel1995} and others is to select the year at which a plot of the cumulative number of events with respect to time becomes (or starts to resemble) a straight line. This is done for different magnitude ranges. Following this approach, we determined the year of completeness for magnitude ranges 3--4, 4--5, 5--6 and 6--7. Our choices are illustrated in Fig.~\ref{fig:completeness}, where we show the normalized cumulative number of events with time for each of the seismic zones and magnitude ranges, and indicate our selection of the year of completeness in each case. The values shown in this figure and those corresponding to the uniform model of the entire region of interest are summarized in Table \ref{tab:completeness}. For events $M_w>7$, in which case there are no sufficient data to determine a proper threshold, we assumed the catalog to be complete from the earliest time at which there is record of large magnitude earthquakes.

@@

We defined these year intervals based on the data magnitude-time distribution (see Fig.~\ref{fig:scatter}) and the idea of distinguishing between historical, early-instrumental, and modern-instrumental catalogs. Having augmented the original catalog with $M>3$ earthquakes beginning in 2000 also influenced our choice of the intervals because at this point the uniformity of the catalog is altered. Last, we also computed $M_c$ values for the complete regional catalogs, but in this case, in order to maintain uniformity throughout time, we only considered $M>4$ events. Fig.~\ref{fig:mc} illustrates the selection of $M_c$ for the case of the main three northern Iran seismic regions, and Table \ref{tab:completeness} presents the final selection of $M_c$ values for these and the other regions, including the uniform northern Iran model for the complete region of interest.

@@

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-06.pdf} 
    \caption{Selection of completeness magnitude ($M_c$) values for the three main seismic regions in northern Iran. The blue squares indicate the computed goodness-of-fit test values and the green circles indicate the cumulative number of events. Both are shown as functions of the earthquake magnitude. The former is plotted in linear scale (left axes) and the latter is plotted in logarithmic scale (right axes). The horizontal dashed lines indicate desired thresholds for the goodness-of-fit values at 80 and 90 percent. The vertical red lines indicate the chosen $M_c$ values for each zone and catalog year interval.}
    \label{fig:mc}
\end{figure*}

### 2016-07-09 conclusions.tex
implemented final comments from Naeem

We present results from a comprehensive seismic hazard analysis for the region of northern Iran based on a smoothed seismicity approach, including regional maps of expected peak ground accelerations for different seismic-zonation models and magnitude thresholds. In particular, we calculated the hazard for the region of interest using two basic models, a five-seismic zone regions model and a uniform seismic zone model. Our calculations yield results similar to those of other studies but offer a more complete picture than previous hazard analysis due to the usage of a more complete catalog and inclusion of low-magnitude events. In our analysis we found it convenient to combine both the regional and uniform models as a mean to consider the epistemic uncertainty intrinsic to the subjective choice of models and seismic zones. Such a combination yielded hazard maps that adjusted better to the visual interpretation of the historical and instrumental seismicity observed in the region of interest, and the to the location of known seismically active faults. While no one model can be deemed more appropriate than others, we believe our results constitute a positive contribution to the understanding of seismic hazard in the region, especially for rural ares which are not the subject of specialized hazard studies such as those conducted for major urban areas. Future work to improve the results presented here should consider alternative attenuation relationships, especially for estimating other ground motion parameters such as response spectral accelerations for different periods and site effects.

>>>>

We present results from a comprehensive seismic hazard analysis for the region of northern Iran based on a smoothed seismicity approach, including regional maps of expected peak ground accelerations for different seismic zonation models and magnitude thresholds. In particular, we calculated the hazard for the region of interest using two basic models, a five-seismic zone regions model and a uniform seismic zone model. Our calculations yield results similar to those of other studies but offer a more complete picture than previous hazard analysis due to the usage of a more complete catalog and inclusion of low-magnitude events. In our analysis we found it convenient to combine both the regional and uniform models as a mean to consider the epistemic uncertainty intrinsic to the subjective choice of models and seismic zones. Such a combination yielded hazard maps that adjusted better to the visual interpretation of the historical and instrumental seismicity observed in the region of interest, and the to the location of known seismically active faults. While no one model can be deemed more appropriate than others, we believe our results constitute a positive contribution to the understanding of seismic hazard in the region, especially for rural ares which are not the subject of specialized hazard studies such as those conducted for major urban areas. Future work to improve the results presented here should consider alternative attenuation relationships, especially for estimating other ground motion parameters such as response spectral accelerations for different periods and site effects.

### 2016-07-09 parameters.tex
implemented final comments from Naeem

Following the general approach described in the previous sections and using the information collected about the seismic catalog and its completeness, we computed the seismicity parameters $a$ and $b$, and the maximum magnitude $M_{\max}$ using the seismic hazard assessment software HA3 (version 3). This software implements the procedures developed by \citet{Kijko_1989_BSSA, Kijko_1992_BSSA}, and subsequent updates by \citet{Kijko_2004_PAG}. HA3 is widely used for this type of computations and is regarded as a reliable tool in seismic hazard analysis \citep[see, for instance,][]{Karimiparidari2013, Khodaverdian_2016_BSSA}.

>>>>

Following the general approach described in the previous sections and using the information collected about the seismic catalog and its completeness, we computed the seismicity parameters $a$ and $b$, and the maximum magnitude $M_{\max}$ using the seismic hazard assessment software HA3, version 3 \citep{Kijko_2004_HA3}. This software implements the procedures developed by \citet{Kijko_1989_BSSA, Kijko_1992_BSSA}, and subsequent updates by \citet{Kijko_2004_PAG}. HA3 is widely used for this type of computations and is regarded as a reliable tool in seismic hazard analysis \citep[see, for instance,][]{Karimiparidari2013, Khodaverdian_2016_BSSA}.

### 2016-07-09 results.tex
implemented final comments from Naeem

An additional consideration to make here is the fact that any one model cannot be uniquely correct. They all carry a certain level of epistemic uncertainty which makes them equally accurate and inaccurate at the same time. A means to address the uncertainty intrinsic to each model is to use combinations of different models. We consider different types of combinations to aid our analysis. First, we combine results at in the R and U models independently by averaging the two magnitude results. These combinations follow the hypothesis that in areas where mixed reinforced and unreinforced buildings are present, the hazard and associated expected ground motions will be an average of both magnitude thresholds. Fig.~\ref{fig:pgaavgs}a shows the result of this combination for the R models and Fig.~\ref{fig:pgaavgs}b shows the result for the combined U models. Second, we consider the combination of the two underlying hazard models R and U for the separate magnitude levels. These results are shown in Fig.~\ref{fig:pgaavgs}c. Last, we consider an overall combination of both hazard models and magnitude thresholds, as shown in Fig.~\ref{fig:pgaavgs}d. At all times we keep separate the probabilities of exceedance.

@@

We also note that the overall distribution of PGA values shown in Fig.~\ref{fig:pgaavgs}c are the ones that best agree with the recent study by \citet{Khodaverdian_2016_BSSA}. These estimates also agree, on average, with other previous studies. For the north-west seismic regions of Azerbaijan and Zagros, for instance, \citet{Tavakoli1999} predicts maximum mean PGA values of 0.45 g near the North Tabriz Fault and the North Tehran Fault zones, and 0.4 g around the city of Urmia, where we find equivalent values to be between 0.4 and 0.6 g, and about 0.35 g, respectively. The results from \citet{Vafaie2011} for the city of Tabriz, on the other hand, are slightly higher than the averages shown in Fig.~\ref{fig:pgaavgs}c and are closer to those we present in Fig.~\ref{fig:pga}b, which are in the order of 0.8 g and 0.6 for the models 2U5.5 and 10U5.5. In the north-central region of Alborz, \citet{Ghodrati2003} and \citet{Boostan2015} predict PGA values of the order 0.27--0.46 g and 0.42--0.48 for Tehran, where we found equivalent results to be between 0.3 and 0.45 g. Similarly, \citet{Abdi2013} reports expected PGA values in the vicinity of Karaj and Tehran to be on the order of 0.3 g. Results from \citet{Ghodrati2008} for the Gilan province around the city of Rasht are also of the same order of those we show in Figs.~\ref{fig:pga} and \ref{fig:pgaavgs}c. Finally, for the Kopeh Dagh region, and in particular, for the city of Bojnurd in the northeast province of Khorasan, \citet{Rahgozar2012} reports expected PGA values for return periods of 475 and 2475 years in range of 0.15--0.22 g and 0.28--0.49 g, respectively, where for magnitudes exceeding $M_w$ 6.5 we found equivalent values to be in the order of 0.2--0.4 and 0.4--0.6 g.

@@

Table \ref{tab:pga} shows a condensed version of these comparisons between our results and those of other previous studies for selected cities in northern Iran. As mentioned before, overall, our results are close or within the ranges from most studies. The only noticeable exceptions are the results of \citet{Golara2014}, whose expected PGA values are significantly higher the rest, especially in the cities of Tabriz, Karaj, Tehran and Mashhad. This is likely due to the authors working with mower magnitude thresholds (not clearly reported) than other studies where hazard is usually estimated based on earthquake magnitudes exceeding $M_w > 5$. Although not included here for brevity we also computed expected PGA values for a magnitude threshold of $M_w$ 4.5 and 5.0. We found the former case to yield results closer to those of \citet{Golara2014}, but not to the other studies. Using lower magnitude thresholds may be of relevance if evaluating poorly constructed and unreinforced buildings vulnerable to low-to-moderate magnitude earthquakes.

>>>>

An additional consideration to make here is the fact that any one model cannot be uniquely correct. They all carry a certain level of epistemic uncertainty which makes them equally accurate and inaccurate at the same time. A means to address the uncertainty intrinsic to each model is to use combinations of different models. We consider different types of combinations to aid our analysis. First, we combine results at in the R and U models independently by averaging the two magnitude results. These combinations follow the hypothesis that in areas where mixed reinforced and unreinforced buildings are present, the hazard and associated expected ground motions will be an average of both magnitude thresholds. Fig.~\ref{fig:pgaavgs}a shows the result of this combination for the R models and Fig.~\ref{fig:pgaavgs}b shows the result for the combined U models. Second, we consider the combination of the two underlying hazard models R and U for the separate magnitude levels. These results are shown in Fig.~\ref{fig:pgaavgs}c. Last, we consider an overall combination of both hazard models and magnitude thresholds, as shown in Fig.~\ref{fig:pgaavgs}d.

@@

We also note that the overall distribution of PGA values shown in Fig.~\ref{fig:pgaavgs}c are the ones that best agree with the recent study by \citet{Khodaverdian_2016_BSSA}. These estimates also agree, on average, with other previous studies. For the north-west seismic regions of Azerbaijan and Zagros, for instance, \citet{Tavakoli1999} predicts maximum mean PGA (10\% in 50 years) values of 0.45 g near the North Tabriz Fault and the North Tehran Fault zones, where we find equivalent values to be between 0.2 and 0.4 g. The results from \citet{Vafaie2011} for the city of Tabriz, on the other hand, are slightly higher than the averages shown in Fig.~\ref{fig:pgaavgs}c and are closer to those we present in Fig.~\ref{fig:pga}b, which are in the order of 0.8 and 0.6 g for the models 2U5.5 and 10U5.5. In the north-central region of Alborz, \citet{Ghodrati2003} and \citet{Boostan2015} predict PGA values of the order 0.27--0.46 g and 0.42--0.48 for Tehran, where we found equivalent results to be between 0.2 and 0.35 g, which are closer to the values reported by \citet{Abdi2013}, who predicts PGA values in the vicinity of Karaj and Tehran to be on the order of 0.3 g. Results from \citet{Ghodrati2008} for the Gilan province around the city of Rasht are also of the same order of those we show in Figs.~\ref{fig:pga} and \ref{fig:pgaavgs}c. Finally, for the Kopeh Dagh region, and in particular, for the city of Bojnurd in the northeast province of Khorasan, \citet{Rahgozar2012} reports expected PGA values for return periods of 475 and 2475 years in range of 0.15--0.22 g and 0.28--0.49 g, respectively, where for magnitudes exceeding $M_w$ 6.5 we found equivalent values to be in the order of 0.2--0.4 and 0.4--0.6 g.

@@

Table \ref{tab:pga} shows a condensed version of these comparisons between our results and those of other previous studies for selected cities in northern Iran. As mentioned before, overall, our results are close or within the ranges from most studies. The only noticeable exceptions are the results of \citet{Golara2014}, whose expected PGA values are significantly higher the rest, especially in the cities of Tabriz, Karaj, Tehran and Mashhad. This is likely due to the authors working with lower magnitude thresholds (not clearly reported) than other studies where hazard is usually estimated based on earthquake magnitudes exceeding $M_w > 5$. Although not included here for brevity, we also computed expected PGA values for magnitude thresholds $M_w$ 4.5 and 5.0. We found the former case to yield results closer to those of \citet{Golara2014}, but not to the other studies. Using lower magnitude thresholds may be of relevance if evaluating poorly constructed and unreinforced buildings vulnerable to low-to-moderate magnitude earthquakes. We should also note that, in general, the smoothed seismicity method used here tends to lead to lower hazard estimations than those obtained using specific fault definitions (as opposed to point sour\-ces).
